Molecular Gas Dynamics of the lensed wet-merger RXSJ1131-1123 at $z$=0.654

### 2016-06-20 WRITEUP/ms.tex
update table w/ notes and formatting
decomposed Mgas for RXJ and companion, updated formatting and rewording based on DR's comments

\subsection{\cco Emission} % DONE
We detect resolved \cco line emission toward RXJ1131. The spectrum is shown in
\Fig{co32spec}, which seems to have a double-peak profile.
This is expected if RXJ1131 is truly a disk galaxy (see previous
sections). The high phase noise in the calibration leads to a low SNR
detection. We thus estimate the line intensity to be
35.7\,$\pm$\,21.9 Jy\,\kms by summing up fluxes over the FWZI
linewidth used to infer \bco line intensity ($\sim$700 \kms).

@@

Assuming the spatial extents between \bco and \cco are similar and therefore
a negligible differential lensing effect, the line intensities
correspond to a brightness temperature ratio of
$r_{\rm 32}$\,=\,T$_{\small \cco}$$/$T$_{\small \bco}$\,=\,0.66\,$\pm$\,0.41.

@@

We fit a power-law spectrum to the
decomposed IRAC photometry to disentangle the observed total flux
at MIPS 24\,$\micron$ into the foreground and
background galaxies. We find a spectral index of $\alpha\sim-$1.8 and
$\alpha\sim-0.85$ for the lensing galaxy and RXJ1131, respectively.
This is consistent with the mean 3.6\,$-$\,8\,$\micron$
spectral slope of
$\alpha$\,=\,$-$1.07\,$\pm$\,0.53 found for unobscured AGN
\citep{Stern05a}. An extrapolation of the fit to 24$\micron$
yields 33.96\,$\pm$\,0.01\,mJy and 25.19\,$\pm$\,0.03\,mJy
for the foreground galaxy and RXJ1131, respectively.
We note that the IRAC photometry includes
emission from old stellar population and is prone to
dust extinction. Hence, the decomposed fluxes are only
our best estimate of the warm dust emission.
We incorporate the decomposed 24$\micron$ point in our
SED fitting to provide some constraints on
the Wien tail beyond the dust peak
of the spectral energy distribution (SED) of RXJ1131.
Details of the SED modeling is presented in \Sec{SED}.

Extraction of the {\it Herschel}/SPIRE photometry was
carried out using \ncode{sussextractor} within the Herschel Interactive
Processing Environment (HIPE; \citealt{Ott10a})
on Level 2 maps obtained from the Herschel Science Archive.
These maps were processed by the SPIRE pipeline
version 13.0 within HIPE. The \ncode{sussextractor} task estimates
the flux density from an image convolved with a kernel
derived from the SPIRE beam. The flux density
measured by \ncode{sussextractor} is additionally confirmed
using the Timeline Fitter, which performs photometry
by fitting a 2D elliptical Gaussian to the Level 1 data at the
source position given by the output of \ncode{sussextractor}.

@@

%--------------------------------------------------------------------------
%                                Analysis
%--------------------------------------------------------------------------
\section{Analysis} \label{sec:anal}
\subsection{Lens modeling} \label{sec:lensmodel} %DONE
At the angular resolution of the \bco data, the images are resolved
but not within the multiply lensed emission (i.e., sub-optimal for
performing lens modeling). Nevertheless, the high spectral
resolution of this data provides dynamical information on
spatial scales smaller than the beam (see \Fig{CO21highO}).
Hence, with the high SNR, we reconstruct the intrinsic gas
dynamics by carrying out a parametric lens modeling over different
channel slices of the interferometric data using our lensing code
\uvmcmcfit (\citealt{uvmcmcfit15a}; see \citealt{Bussmann15a} for details of
the code). Model of each slice in \Fig{model} thus provides
properties on the intrinsic kinematics. The slices
are obtained by binning the data over the full linewidth
$\sim$700\,\kms by five channels,
that is, into seven slices, to increase the SNR.

@@

We first obtain a preliminary lens model for each channel slice independently,
where their lens parameters are allowed to vary and are initialized according
to the aforementioned way. We obtain the final model
by repeating the modeling over each slice but fixing their lens parameters
according to the overall median in the preliminary models,
as listed in \Tab{lens}.
This ensures that all models share the same lens profile.
The magnification factors in \Tab{model} are determined by taking the ratio
between the image plane flux and the source plane flux of each model.

@@

\subsubsection{Interpretation and Caveats} \label{sec:caveat} %DONE
The reconstructed source locations in \Fig{model} demonstrate
an intrinsic velocity gradient across the source plane, which is
indicative of a disk-like galaxy.
Additional support to the disk conjecture %conjecture of a disk morphology
can be found in the double-horned line profile (\Fig{CO21spec})
and the observed (image plane) velocity field (\Fig{CO21highO}). Furthermore,
C06 also find that the reconstructed source plane emission in optical-NIR
is best-reproduced using a $n$\,=\,1 Sersic profile.
Our interpretation therefore tends in favor of a disk morphology for RXJ1131.

% the companion
One other interesting result from our lens model is that a better fit is
found for the red-most channel if we add a second source component (see
top left panel in \Fig{model}). This is consistent with previous results
reported by \citet{Brewer08a}, who find an optically faint companion
(component F in their paper) $\sim$2.4\,kpc away from the AGN host galaxy,
and C06, who find evidence for an interacting galaxy near RXJ1131.
Spatially, the red velocity component of the CO also coincides with this
component F. While we cannot quantify the type of
merger (major v.s. minor) with a mass ratio given the
spatial resolution, the symmetric velocity gradient retained under
such close separation between the two lean support
to the idea of a minor merger.
% For major mergers, two galaxy mass centers separated on the same scale as the
% gas disk (4 kpc or more), the gas dynamics would be strongly perturbed and
% would deviate substantially from that of a simple disk.
% In a minor merger (a mass ratio of more than 3:1) the more massive galaxy, including its gas distribution, can retain a significant angular momentum and resemble a rotating disk.

The spatial resolution of the data in hand
is a few arcsec, which implies that despite the high SNR and spectral
resolution, this data is insufficient to constrain the
intrinsic sizes of the lensed galaxies, and thus the magnification
factors may be under-predicted.
% magnification factor is underpredicted, see Bussmann15

\begin{figure*}[tbph]
\centering
\begin{tabular}{c}
\includegraphics[trim=0 0 0 0, clip, width=1.0\textwidth]{../Figures/PostageStampModel.eps} \\
\includegraphics[width=0.98\textwidth]{../Figures/PostageStampResiduals.eps}
\end{tabular}
\caption{Each panel corresponds to a lens model of RXJ1131 performed over a
channel slice $\sim$100\,\kms of the \bco data. Top: channel maps of the
PdBI \bco emission (red) overlaid on our best-fit lens models (grayscale).
The location of the foreground lensing galaxy is indicated by a black dot and
its critical curve is traced by the orange solid line. The locations and
morphologies (half-light radii) of the reconstructed sources are
represented by magenta ellipses (see text in \Sec{caveat} for caveats).
The caustic curves are represented as cyan lines. The beam of the
PdBI observations is shown in the bottom right corner of each panel.
Bottom: residual images of the best-fit models, obtained by
taking the Fourier transform after subtracting the best-fit model from the
data in the $uv$-domain. Contours start
at $\pm$3$\sigma$ and increment at steps of 3$\times$2$^n\sigma$,
where $n$ is a positive integer.
\label{fig:model}}
\end{figure*}

\subsubsection{Differential Lensing} \label{sec:differential} %DONE
\input{table_magFactor}
Previous lens models of RXJ1131 find evidence for
differential lensing across {\it HST}
$H$-, $V$-, and $I$-band (C06), where the
magnification factor ranges from 10.9 to 7.8. C06 attribute
the decrease toward the IR regime is caused by the more extended
IR emission. The highly asymmetric \bco line profile suggests that
differential lensing is also non-negligible for CO, where the redshifted
emission is apparently much brighter than that in the
blueshifted part, consistent
with the source positions relative to the caustics in \Fig{model}.
The red wing emission mainly originates near the cusp
in the caustic, whereas the blue wing emission is located beyond the caustics.
The magnification factor for CO ranges from 8.7 to 3.1 (\Tab{model}).

@@

\subsubsection{Gas Mass}
To derive the molecular gas mass, we first correct for lensing magnification
using the respective magnification factors listed in \Tab{model}, which to
first order takes into account effect of differential lensing. This yields
$I_{\small \bco}$\,=\,5.9$\pm$2.7 Jy\,\kms,
where the uncertainty includes those on
the magnification factors.
We note that this includes emission from the companion
as the spatial resolution of our data do not allow us decompose its
contribution from the total line flux.

Without direct \rot{1}{0} measurements,
we derive \rot{1}{0} intensity
using a brightness temperature ratio of $r_{\rm 32}$\,=\,1, which was chosen
solely to facilitate a comparison with other U/LIRGs.
For the same reason, we
adopt a luminosity-to-mass conversion factor of
\alphaco\,=\,0.8 (K \kms pc$^2$)\pmOne, this yields a
CO luminosity \Lp\,=\,(3.47\,$\pm$\,1.58)\E{10}\,\LpU and a gas mass
$M_{\rm gas}$ of (2.8$\pm$1.3)\,\E{10} \Msun. This is likely
an overestimation for RXJ1131 due to a contribution from the companion.
The inferred gas mass is comparable to
C13 sample ($M_{\rm gas}$\,=\,1.5\E{10}\Msun) as well as
local U/LIRGs \citep{Solomon97a, Sanders96a} but higher than
local LIRGs (U14).

>>>>

\subsection{\cco Emission}
We detect \cco line emission toward RXJ1131 at {\bf BLAH}$\sigma$ significance.
The spectrum is shown in \Fig{co32spec}, which seems to have a double-peak profile.
This is expected if RXJ1131 is truly a disk galaxy (see previous
sections). The high phase noise in the calibration leads to a low SNR
detection. We thus estimate the line intensity to be
35.7\,$\pm$\,21.9 Jy\,\kms by summing up fluxes over the FWZI
linewidth used to infer \bco line intensity ($\sim$700 \kms).

@@

Assuming the spatial extents between \bco and \cco are similar and therefore
magnified by similar amounts, the line intensities
correspond to a brightness temperature ratio of
$r_{\rm 32}$\,=\,T$_{\small \cco}$$/$T$_{\small \bco}$\,=\,0.66\,$\pm$\,0.41.

@@

We fit a power-law spectrum to the
decomposed IRAC photometry to disentangle the observed total flux
at MIPS 24\,$\micron$ into the foreground and
background galaxies. We find a spectral index of $\alpha\sim-$1.8 and
$\alpha\sim-0.85$ for the lensing galaxy and RXJ1131, respectively.
This is consistent with the mean 3.6\,$-$\,8\,$\micron$
spectral slope of
$\alpha$\,=\,$-$1.07\,$\pm$\,0.53 found for unobscured AGN
\citep{Stern05a}. An extrapolation of the fit to 24$\micron$
yields 33.96\,$\pm$\,0.01\,mJy and 25.19\,$\pm$\,0.03\,mJy
for the foreground galaxy and RXJ1131, respectively.
We note that the IRAC photometry includes
emission from old stellar populations and is prone to
dust extinction. Hence, the decomposed fluxes are only
our best estimate of the warm dust emission.
We incorporate the decomposed 24$\micron$ point in our
SED fitting to provide some constraints on
the Wien tail beyond the dust peak
of the spectral energy distribution (SED) of RXJ1131.
Details of the SED modeling is presented in \Sec{SED}.

Extraction of the {\it Herschel}/SPIRE photometry was
carried out using \ncode{sussextractor} within the Herschel Interactive
Processing Environment (HIPE; \citealt{Ott10a})
on Level 2 maps obtained from the Herschel Science Archive.
These maps were processed by the SPIRE pipeline
version 13.0 within HIPE. The \ncode{sussextractor} task estimates
the flux density from an image convolved with a kernel
derived from the SPIRE beam. The flux density
measured by \ncode{sussextractor} is additionally confirmed
using the Timeline Fitter, which performs photometry
by fitting a 2D elliptical Gaussian to the Level 1 data at the
source position given by the output of \ncode{sussextractor}. The fluxes
obtained from both methods are consistent within the uncertainties.

@@

%--------------------------------------------------------------------------
%                                Analysis
%--------------------------------------------------------------------------
\section{Analysis} \label{sec:anal}
\subsection{Lens modeling} \label{sec:lensmodel}
At the angular resolution of the \bco data, the images are resolved over
$\sim$2 resolution elements. Given the extent of the lensed emission,
this implies that our \bco observations do not resolve
structures (e.g. knots and arcs) of the lensed emission
(i.e., sub-optimal for
performing lens modeling). Nevertheless, the high spectral
resolution of this data provides dynamical information on
spatial scales smaller than the beam (see \Fig{CO21highO}).
Hence, with the high SNR, we reconstruct the intrinsic gas
dynamics by carrying out a parametric lens modeling over different
channel slices of the interferometric data using our lensing code
\uvmcmcfit (\citealt{uvmcmcfit15a}; see \citealt{Bussmann15a} for details of
the code). Models of each slice thus provide the
properties on the intrinsic kinematics. To increase the SNR for modeling, the
slices are obtained by binning five native channels
($\Delta v$\,$\sim$\,21.5\kms) over the full
linewidth of $\sim$750\,\kms in the original data, resulting in seven
independent channels.

@@

We first obtain a preliminary lens model for each channel slice independently,
where their lens parameters are allowed to vary and are initialized according
to the aforementioned way. We obtain the final model
by repeating the modeling over each slice but fixing their lens parameters
to the overall median in the preliminary models,
as listed in \Tab{lens}.
This ensures that all models share the same lens profile.
The magnification factors in \Tab{model} are determined by taking the ratio
between the image plane flux and the source plane flux of each model.

@@

\subsubsection{Interpretation of the source-plane morphology} \label{sec:caveat}
The reconstructed source locations in \Fig{model} demonstrate
an intrinsic velocity gradient across the source plane, which is
indicative of a disk-like galaxy.
Additional support to the disk conjecture
can be found in the double-horned line profile (\Fig{CO21spec})
and the observed (image plane) velocity field (\Fig{CO21highO}). Furthermore,
C06 also find that the reconstructed source plane emission in optical-NIR
is best-reproduced using a $n$\,=\,1 Sersic profile.
We thus interpret RXJ1131 to be a disk galaxy.

% the companion
One other interesting result from our lens model is that a better fit is
found for the red-most channel if we add a second source component (see
top left panel in \Fig{model}). This is consistent with previous results
reported by \citet{Brewer08a}, who find an optically faint companion
(component F in their paper) $\sim$2.4\,kpc away from the AGN host galaxy,
and C06, who find evidence for an interacting galaxy near RXJ1131.
Spatially, the red velocity component of the CO emission
also coincides with this component F. It is therefore evident that we
detect \bco emission in the companion galaxy.
To quantify the type of merger (major v.s. minor) conventionally with a mass
ratio, we decompose the total line flux into two components:
one from RXJ1131 and the other from its companion.
Since the companion is only detected in the red-most channel, we
use the best-fit flux densities and magnification factors obtained
from the model of this channel to derive the gas mass in the companion.
Without direct \rot{1}{0} measurements, we derive \rot{1}{0} intensity
assuming a brightness temperature ratio of $r_{\rm 21}$\,=\,1, which was chosen
solely to facilitate a comparison with other U/LIRGs. % also ULIRGs are found to be thermalized up to Jup = 3 or 4
For the same reason, we adopt a luminosity-to-mass conversion factor of
\alphaco\,=\,0.8 (K \kms pc$^2$)\pmOne. This yields a gas mass
$M_{\rm gas}$ of (1.68\pmm0.68)\,\E{10} \Msun. 
%statistical unc. only (i.e. draw from from distributions of on mag. factor & intrinsic flux)
For the molecular gas mass in RXJ1131, we derive
its intrinsic line flux over the FWZI linewidth
using the respective magnification
factors listed in \Tab{model}, which to
first order takes into account effect of differential lensing.
This yields $I_{\small \bco}$\,=\,5.41$\pm$2.71 Jy\,\kms,
where the uncertainty includes those on
the magnification factors.
Adopting the same brightness temperature ratio and \alphaco\ as
used for the companion, this corresponds to a gas mass of
$M_{\rm gas}$\,=\,(3.17\pmm1.59)\,\E{10} \Msun.
The implies a gas mass ratio of $\sim$2:1.
Therefore, the system is a gas-rich major merger according to
the classification scheme commonly used in literature.

The spatial resolution of the data in hand
is a few arcsec, which implies that despite the high SNR and spectral
resolution, this data is insufficient to constrain the
intrinsic sizes of the lensed galaxies, and thus the magnification
factors may be under-predicted. % see Bussmann15

\begin{figure*}[tbph]
\centering
\begin{tabular}{c}
\includegraphics[trim=0 0 0 0, clip, width=1.0\textwidth]{../Figures/PostageStampModel.eps} \\
\includegraphics[width=0.98\textwidth]{../Figures/PostageStampResiduals.eps}
\end{tabular}
\caption{Each panel corresponds to a lens model of RXJ1131 performed over a
channel slice $\sim$100\,\kms of the \bco data. Top: channel maps of the
PdBI \bco emission (red) overlaid on our best-fit lens models (grayscale).
The location of the foreground lensing galaxy is indicated by a black dot and
its critical curve is traced by the orange solid line. The locations and
morphologies (half-light radii) of the reconstructed sources are
represented by magenta ellipses.
The caustic curves are represented as cyan lines. The beam of the
PdBI observations is shown in the bottom right corner of each panel.
Bottom: residual images of the best-fit models, obtained by
taking the Fourier transform after subtracting the best-fit model from the
data in the $uv$-domain. Contours start
at $\pm$3$\sigma$ and increment at steps of 3$\times$2$^n\sigma$,
where $n$ is a positive integer.
\label{fig:model}}
\end{figure*}

\subsubsection{Differential Lensing} \label{sec:differential} %DONE
\input{table_magFactor}
Similar to previous studies of RXJ1131, where
differential lensing across {\it HST}
$H$-, $V$-, and $I$-band has been detected with a
magnification factor decreasing from 10.9 to 7.8 (C06),
the highly asymmetric \bco line profile suggests that
differential lensing is also non-negligible for the CO emission, whereby
causing the redshifted emission to be apparently much brighter than the
blueshifted component.
Such an effect is consistent with the source plane
positions relative to the caustics in \Fig{model}; the red wing emission mainly originates near the cusp
in the caustic, whereas the blue wing emission is located beyond the caustics.
From our best-fit lens models, the magnification factor varies from 8.7 to 3.1 across the \bco line (\Tab{model}).

%If we estimate the intrinsic I:
%red I =
%blue wing I =

@@

\subsubsection{Gas Mass}
To derive the molecular gas mass, we first correct for lensing magnification
using the respective magnification factors listed in \Tab{model}, which to
first order takes into account effect of differential lensing. This yields
$I_{\small \bco}$\,=\,5.9$\pm$2.7 Jy\,\kms,
where the uncertainty includes those on
the magnification factors.

Without direct \rot{1}{0} measurements,
we derive \rot{1}{0} intensity
using a brightness temperature ratio of $r_{\rm 32}$\,=\,1, which was chosen
solely to facilitate a comparison with other U/LIRGs.
For the same reason, we
adopt a luminosity-to-mass conversion factor of
\alphaco\,=\,0.8 (K \kms pc$^2$)\pmOne, this yields a
CO luminosity \Lp\,=\,(3.47\,$\pm$\,1.58)\E{10}\,\LpU and a gas mass
$M_{\rm gas}$ of (xx\pmm xx)\,\E{10} \Msun. 
BLAH
The inferred gas mass is comparable to
C13 sample ($M_{\rm gas}$\,=\,1.5\E{10}\Msun) as well as
local U/LIRGs \citep{Solomon97a, Sanders96a} but higher than
local LIRGs (U14).

### 2016-06-20 WRITEUP/table_medianLens.tex
decomposed Mgas for RXJ and companion, updated formatting and rewording based on DR's comments

\begin{deluxetable}{lcc}[!htbp]
\tabletypesize{\scriptsize}
\tablecolumns{3}
\tablecaption{Lens parameters from preliminary models}
\tablehead{
\multicolumn{2}{c}{Parameters} &
\colhead{Median values} % weighted-mean of median of each
}
\startdata
Offset in RA    & (\arcsec)   &  0.004$\pm$0.027\\
Offset in Dec    & (\arcsec)   & 0.003$\pm$0.027\\
Axial Ratio      &             & 0.56$\pm$0.16\\
Position Angle   & (deg)       & 103$\pm$22\\
Einstein Radius  & (\arcsec)   & 1.833$\pm$0.002
\enddata
\label{tab:lens}
\tablecomments{Parameters describing the foreground lens are
obtained based on the preliminary models (see text for details).
All angular offsets are with respect to
$\alpha$\,=\,11$^{\rm h}$31$^{\rm m}$51\fs44,
$\delta$\,=\,-12\degr31\arcmin58\farcs3 (J2000).}
\end{deluxetable}

>>>>

\begin{deluxetable}{lcc}[!htbp]
\tabletypesize{\scriptsize}
\tablecolumns{3}
\tablecaption{Lens parameters constrained by models of seven velocity channels}
\tablehead{
\multicolumn{2}{c}{Parameters} &
\colhead{Median values} % weighted-mean of median of each
}
\startdata
Offset in RA    & (\arcsec)   &  0.004$\pm$0.027\\
Offset in Dec    & (\arcsec)   & 0.003$\pm$0.027\\
Axial Ratio      &             & 0.56$\pm$0.16\\
Position Angle   & (deg)       & 103$\pm$22\\
Einstein Radius  & (\arcsec)   & 1.833$\pm$0.002
\enddata
\label{tab:lens}
\tablecomments{Parameters describing the foreground lens are
obtained based on the median in the preliminary models (see text for details).
All angular offsets are with respect to
$\alpha$\,=\,11$^{\rm h}$31$^{\rm m}$51\fs44,
$\delta$\,=\,$-$12\degr31\arcmin58\farcs3 (J2000).}
\end{deluxetable}
